---
title: "Haralson County, GA: Insurance Status & Poverty by Age Group"
author: "Danielle McMillion"
date: "`r Sys.Date()`"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
library(shiny)
library(here)

# Load cleaned dataset
data <- read_csv(here("Haralson_County_Final", "data", "haralson_data_cleaned.csv"))

# Set correct age group factor levels
age_levels <- c("18 and under", "19-34", "35-64", "65+")
data$Age_Group <- factor(data$Age_Group, levels = age_levels)
```

Key Takeaways: 
Based on data collected by the U.S. Census, Haralson County, Georgia has a high population of their younger residents that live below the Federal Poverty line. Additionally, a high percentage of their residents that are publically insured are among the elderly population. 

Row
-----------------------------------------------------------------------
### Poverty Rate by Age Group
```{r}
plot_ly(data = data %>% filter(!is.na(Poverty_Percent)),
        x = ~Age_Group, y = ~Poverty_Percent, type = 'bar',
        hovertext = ~paste("Poverty Rate:", round(Poverty_Percent, 1), "%\nCount (individuals):", Poverty_Count),
        hoverinfo = "text",
        marker = list(color = 'mediumpurple4')) %>%
  layout(title = "Haralson County, GA: Poverty Rate (%)",
         xaxis = list(title = "Age Group (yrs)"),
         yaxis = list(title = "% in Poverty (100% FPL)"))
```

Row
-----------------------------------------------------------------------
### Insurance Coverage by Age Group with Legend Toggle

```{r}
plot_ly() %>%
  add_trace(data = data, x = ~Age_Group, y = ~Private_Percent,
            name = "Private Insurance", type = 'bar',
            hovertext = ~paste("Private Insurance:", round(Private_Percent, 1), "%<br>Count:", Private_Count),
            hoverinfo = "text",
            marker = list(color = 'steelblue')) %>%
  add_trace(data = data, x = ~Age_Group, y = ~Public_Percent,
            name = "Public Insurance", type = 'bar',
            hovertext = ~paste("Public Insurance:", round(Public_Percent, 1), "%<br>Count:", Public_Count),
            hoverinfo = "text",
            marker = list(color = 'seagreen')) %>%
  add_trace(data = data, x = ~Age_Group, y = ~Uninsured_Percent,
            name = "Uninsured", type = 'bar',
            hovertext = ~paste("Uninsured:", round(Uninsured_Percent, 1), "%<br>Count:", Uninsured_Count),
            hoverinfo = "text",
            marker = list(color = 'orangered4')) %>%
  layout(title = "Haralson County, GA: Insurance Coverage by Age Group",
         barmode = 'group',
         xaxis = list(title = "Age Group (yrs)"),
         yaxis = list(title = "Coverage (%)"))
```
Data Description {.tabset}
-----------------------------------------------------------------------

### Dataset Information

- **Source**: U.S. Census Bureau (American Community Survey 2023)
> *Note: The data used is publicly available from the following links:*

- **S1701**: https://data.census.gov/table/ACSST5Y2023.S1701?g=050XX00US13143&tp=false
- **S2701**: https://data.census.gov/table/ACSST5Y2023.S2701?q=Healthy+Insurance+status&g=050XX00US13143 
- **S2703**: https://data.census.gov/table/ACSST5Y2023.S2703?q=Health+Insurance+type&g=050XX00US13143 
- **S2704**: https://data.census.gov/table/ACSST5Y2023.S2704?q=Health+Insurance&g=050XX00US13143 
- **Sample Size**: ~30,000 residents of Haralson County, GA
- **Collection Method**: Aggregated survey estimates from ACS 5-Year Data
- **Population**: Civilian non-institutionalized residents of Haralson County, GA
- **Time Period**: Data collected in 2023

### Summary

This dashboard identifies age-related disparities in insurance coverage and poverty in Haralson County, Georgia. These insights support funding justifications for Federally Qualified Health Centers that target these undeserved groups. 

